---
title: "cpsc471_plots"
output: pdf_document
date: "2024-05-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE) # care

library(tidyverse)
library(dplyr)
library(ggplot2)

# --- for plotting and stuff --- #
# devtools::install_github("bmacGTPM/pubtheme")
library(pubtheme)
library(hrbrthemes)
library(viridis)

```

# File for generating plots and visualizations 

## Import processed data (from local; scp-ed from Grace)

```{r import data}
# import raw data
source_processed_monomer_output = read.csv("../processed_deletion_perturb_out/processed_monomer_output.csv")
source_processed_multimer_output = read.csv("../processed_deletion_perturb_out/processed_multimer_output.csv")

source_processed_AA_monomer_output = read.csv("../processed_deletion_perturb_out/processed_AA_monomer_output.csv")
source_processed_AA_multimer_output = read.csv("../processed_deletion_perturb_out/processed_AA_multimer_output.csv")
```

### Add new variables if needed

```{r visualization preparation function}
# takes in a processed df and a boolean indicating whether it's monomer or multimer
viz_prep = function(processed_df, is_multimer) {
  
  if (!is_multimer) {
    df = processed_df %>%
      # create new possibly useful stats
      mutate(delta_mean_pLDDT_score_norm = delta_mean_pLDDT_score/100.0,
             delta_mean_pLDDT_score_relu = ifelse(delta_mean_pLDDT_score > 0, delta_mean_pLDDT_score, 0),
             delta_pTM_score_relu = ifelse(delta_pTM_score > 0, delta_pTM_score, 0),
             ) %>%
      
      # group stuff
      group_by(seq) %>%
      mutate(delta_mean_pLDDT_score_norm_minmax = (delta_mean_pLDDT_score - min(delta_mean_pLDDT_score))/(max(delta_mean_pLDDT_score) - min(delta_mean_pLDDT_score)),
             delta_pTM_score_norm_minmax = (delta_pTM_score - min(delta_pTM_score))/(max(delta_pTM_score) - min(delta_pTM_score)),
             ) %>%
      ungroup()
  }
  else {
    df = processed_df %>%
      # create new possibly useful stats
      mutate(delta_mean_pLDDT_score_norm = delta_mean_pLDDT_score/100.0,
             delta_mean_pLDDT_score_relu = ifelse(delta_mean_pLDDT_score > 0, delta_mean_pLDDT_score, 0),
             delta_pTM_score_relu = ifelse(delta_pTM_score > 0, delta_pTM_score, 0),
             delta_ipTM_score_relu = ifelse(delta_ipTM_score > 0, delta_ipTM_score, 0),
             ) %>%
      
      # group stuff
      group_by(complex_id) %>%
      mutate(delta_mean_pLDDT_score_norm_minmax = (delta_mean_pLDDT_score - min(delta_mean_pLDDT_score))/(max(delta_mean_pLDDT_score) - min(delta_mean_pLDDT_score)),
             delta_pTM_score_norm_minmax = (delta_pTM_score - min(delta_pTM_score))/(max(delta_pTM_score) - min(delta_pTM_score)),
             delta_ipTM_score_norm_minmax = (delta_ipTM_score - min(delta_ipTM_score))/(max(delta_ipTM_score) - min(delta_ipTM_score)),
             ) %>%
      ungroup()
  }
  
}

```

```{r test viz_prep}
processed_monomer_output = viz_prep(source_processed_monomer_output, is_multimer=FALSE)
processed_multimer_output = viz_prep(source_processed_multimer_output, is_multimer=TRUE)

processed_AA_monomer_output = viz_prep(source_processed_AA_monomer_output, is_multimer=FALSE)
processed_AA_multimer_output = viz_prep(source_processed_AA_multimer_output, is_multimer=TRUE)

```


### Create simplified dataframes 

```{r functions to simplify processed dfs}
simplify_monomer_df = function(df) {
  simple_df = df %>%
  select(seq, delete_index, delta_pTM_score, delta_mean_pLDDT_score) %>%
  rename(mon_delta_pTM_score = delta_pTM_score, mon_delta_mean_pLDDT_score = delta_mean_pLDDT_score)
}

simplify_multimer_df = function(df) {
  # grab homo ones
  temp_homo_ones = df %>%
    filter(which_seq == 1,
           seq1 == seq2) %>%
    select(-seq2) %>%
    rename(seq = seq1) %>%
    select(seq, delete_index, delta_pTM_score, delta_mean_pLDDT_score, delta_ipTM_score)
  
  # grab hetero ones
  temp_hetero_ones = df %>%
    filter(which_seq == 1,
           seq1 != seq2) %>%
    select(-seq2) %>%
    rename(seq = seq1) %>%
    select(seq, delete_index, delta_pTM_score, delta_mean_pLDDT_score, delta_ipTM_score)
  
  # grab hetero twos
  temp_hetero_twos = df %>%
    filter(which_seq == 2,
           seq1 != seq2) %>%
    select(-seq1) %>%
    rename(seq = seq2) %>%
    select(seq, delete_index, delta_pTM_score, delta_mean_pLDDT_score, delta_ipTM_score)
  
  # grab zeros
  temp_zeros = df %>%
    filter(which_seq == 0) %>%
    pivot_longer(cols = c(seq1, seq2), values_to = "seq") %>%
    distinct(seq, .keep_all = TRUE) %>%
    select(seq, delete_index, delta_pTM_score, delta_mean_pLDDT_score, delta_ipTM_score)
  
  # rbind into one big dataframe
  simple_df = rbind(temp_homo_ones, temp_hetero_ones, temp_hetero_twos, temp_zeros) %>%
    rename(multi_delta_pTM_score = delta_pTM_score, 
           multi_delta_mean_pLDDT_score = delta_mean_pLDDT_score,
           multi_delta_ipTM_score = delta_ipTM_score)
  
  simple_df
}



```

```{r simplify processed outputs}
simple_monomer_output = simplify_monomer_df(processed_monomer_output)
simple_multimer_output = simplify_multimer_df(processed_multimer_output) 

simple_regular_output = merge(simple_monomer_output, simple_multimer_output) %>%
  mutate(is_AA = FALSE)

simple_AA_monomer_output = simplify_monomer_df(processed_AA_monomer_output) 
simple_AA_multimer_output = simplify_multimer_df(processed_AA_multimer_output) 

simple_AA_output = merge(simple_AA_monomer_output, simple_AA_multimer_output) %>%
  mutate(is_AA = TRUE)

master_simple_data = rbind(simple_regular_output, simple_AA_output)
master_simple_data_longer = master_simple_data %>%
  pivot_longer(cols = c(mon_delta_pTM_score, mon_delta_mean_pLDDT_score, 
                        multi_delta_pTM_score, multi_delta_mean_pLDDT_score, multi_delta_ipTM_score),
               names_to = "score_type",
               values_to = "delta_score")


```

```{r map sequence to is AA}
seq_is_AA = master_simple_data %>%
  select(seq, is_AA) %>%
  distinct()

```

## Exploratory Data Analysis

### Distributions of scores by metric and is_AA

```{r distributions}
ggplot(data=master_simple_data, aes(mon_delta_mean_pLDDT_score, fill = is_AA)) +
  geom_density(alpha = 0.4) + 
  labs(title = "Monomer delta mean pLDDT score distribution") +
  theme_pub(type="hist")

ggplot(data=master_simple_data, aes(mon_delta_pTM_score, fill = is_AA)) +
  geom_density(alpha = 0.4) + 
  labs(title = "Monomer delta pTM score distribution") +
  theme_pub(type="hist")

ggplot(data=master_simple_data, aes(multi_delta_mean_pLDDT_score, fill = is_AA)) +
  geom_density(alpha = 0.4) + 
  labs(title = "Multimer delta mean pLDDT score distribution") +
  theme_pub(type="hist")

g_multi_delta_pTM = ggplot(data=master_simple_data, aes(multi_delta_pTM_score, fill = is_AA)) +
  geom_density(alpha = 0.4) + 
  labs(title = "Multimer delta pTM score distribution") +
  theme_pub(type="hist")
g_multi_delta_pTM

# ggsave(plot = g_multi_delta_pTM, width = 9, height = 6, dpi = 300, filename = "plots/g_multi_delta_pTM.png")

ggplot(data=master_simple_data, aes(multi_delta_ipTM_score, fill = is_AA)) +
  geom_density(alpha = 0.4) + 
  labs(title = "Multimer delta ipTM score distribution") +
  theme_pub(type="hist")
  

```

## Correlations of scores

### Correlation between different metrics

#### First, plot some delta scores

```{r random delta scatterplot}
g_corr_scatter_ptm_iptm = ggplot(data=master_simple_data, aes(x=multi_delta_pTM_score, y=multi_delta_ipTM_score)) +
  geom_point(aes(color=is_AA), size=2, alpha=0.5) +
  labs(title = "Scatterplot of two delta scores",
       subtitle = "Multimer pTM and multimer ipTM") +
  xlab("delta pTM score") +
  ylab("delta ipTM score") +
  theme_pub(type="scatter")

g_corr_scatter_ptm_iptm

# - save - #
# ggsave(plot = g_corr_scatter_ptm_iptm, width = 9, height = 6, dpi = 300, filename = "plots/g_corr_scatter_ptm_iptm.png")


```


#### Delta score correlations

```{r}
mon_plddt_ptm_corrs = master_simple_data %>%
  group_by(seq) %>%
  summarise(corr = cor(mon_delta_mean_pLDDT_score, mon_delta_pTM_score)) %>%
  left_join(seq_is_AA, by="seq") %>%
  mutate(metric_pair = "mon_pLDDT_pTM")

multi_plddt_ptm_corrs = master_simple_data %>%
  group_by(seq) %>%
  summarise(corr = cor(multi_delta_mean_pLDDT_score, multi_delta_pTM_score)) %>%
  left_join(seq_is_AA, by="seq") %>%
  mutate(metric_pair = "multi_pLDDT_pTM")

multi_ptm_iptm_corrs = master_simple_data %>%
  group_by(seq) %>%
  summarise(corr = cor(multi_delta_pTM_score, multi_delta_ipTM_score)) %>%
  left_join(seq_is_AA, by="seq") %>%
  mutate(metric_pair = "multi_pTM_ipTM")

multi_plddt_iptm_corrs = master_simple_data %>%
  group_by(seq) %>%
  summarise(corr = cor(multi_delta_mean_pLDDT_score, multi_delta_ipTM_score)) %>%
  left_join(seq_is_AA, by="seq") %>%
  mutate(metric_pair = "multi_pLDDT_ipTM")

metric_pair_corrs = rbind(mon_plddt_ptm_corrs, multi_plddt_ptm_corrs, 
                          multi_ptm_iptm_corrs, multi_plddt_iptm_corrs)
  
g_delta_corrs = ggplot(metric_pair_corrs, aes(x=is_AA, y=corr, fill=is_AA)) + 
  geom_violin() +
  geom_point(aes(x=is_AA, y=corr), position="jitter", size = 0.5) +
  facet_wrap(~metric_pair, ncol=4) +
  theme_pub(type="hist") +
  labs(title = "Histogram of Correlations of Delta Scores",
       subtitle = "Within monomer or multimer") 

g_delta_corrs

# - save image - #
# ggsave(plot = g_delta_corrs, width = 9, height = 6, dpi = 300, filename = "../analysis_plots/within_metric_delta_corrs.png")


```



#### [DEPR - single] Multimer pTM and ipTM

```{r}
multi_ptm_iptm_corrs = master_simple_data %>%
  group_by(seq) %>%
  summarise(corr = cor(multi_delta_pTM_score, multi_delta_ipTM_score)) %>%
  left_join(seq_is_AA, by="seq")
  
ggplot(multi_ptm_iptm_corrs, aes(x=is_AA, y=corr, fill=is_AA)) + 
  geom_violin() +
  geom_point(aes(x=is_AA, y=corr), position="jitter", size = 1) +
  theme_pub(type="hist") +
  labs(title = "Histogram of Correlations of Delta Statistics",
       subtitle = "Multimer pTM and ipTM") 

```



### Correlation between monomer and multimer

May find that antibody-antigen scores are much more correlated (Luke's qualitative observations). 

```{r}
cross_plddt_corrs = master_simple_data %>%
  group_by(seq) %>%
  summarise(corr = cor(mon_delta_mean_pLDDT_score, multi_delta_mean_pLDDT_score)) %>%
  left_join(seq_is_AA, by="seq") %>%
  mutate(metric_pair = "pLDDT")

cross_ptm_corrs = master_simple_data %>%
  group_by(seq) %>%
  summarise(corr = cor(mon_delta_pTM_score, multi_delta_pTM_score)) %>%
  left_join(seq_is_AA, by="seq") %>%
  mutate(metric_pair = "pTM")

cross_ptm_iptm_corrs = master_simple_data %>%
  group_by(seq) %>%
  summarise(corr = cor(mon_delta_pTM_score, multi_delta_ipTM_score)) %>%
  left_join(seq_is_AA, by="seq") %>%
  mutate(metric_pair = "pTM_ipTM")

cross_metric_corrs = rbind(cross_plddt_corrs, cross_ptm_corrs, cross_ptm_iptm_corrs)

g_cross_delta_corrs = ggplot(cross_metric_corrs, aes(x=is_AA, y=corr, fill=is_AA)) + 
  geom_violin() +
  geom_point(aes(x=is_AA, y=corr), position="jitter", size = 0.5) +
  facet_wrap(~metric_pair, ncol=4) +
  theme_pub(type="hist") +
  labs(title = "Histogram of Correlations of Delta Scores",
       subtitle = "Between monomer and multimer") 

g_cross_delta_corrs

# - save - #
# ggsave(plot = g_cross_delta_corrs, width = 9, height = 6, dpi = 300, filename = "../analysis_plots/cross_metric_delta_corrs.png")


```

## Heat maps (per sequence)

#### [DEPR] single monomer sequence

```{r per sequence heatmap monomer}
# pick from here
unique_monomer_sequences = unique(processed_monomer_output$seq)
plot_seq = unique_monomer_sequences[1] # can customize

plot_seq_df = processed_monomer_output %>%
  filter(seq == plot_seq,
         delete_index >= 0) %>%
  ungroup() %>%
  arrange(delete_index) %>%
  select(delete_index, delta_mean_pLDDT_score_norm, delta_pTM_score) %>%
  column_to_rownames(var="delete_index")

# plot 1: using absolute normalization (theoretical values in [0,1]; data unchanged)
heatmap(t(as.matrix(plot_seq_df)), Colv = NA, Rowv = NA, scale="none",
        main = "Monomer Delta Values (Un-normalized)",
        xlab = "Residue Index",
        ylab = "Delta Statistic")

# plot 2: using minmax normalization (actual values normalized to be in [0,1])
heatmap(plot_seq_data_matrix, Colv = NA, Rowv = NA, scale="column")

```

#### Single (hetero)multimer sequence

```{r}
# run logic
unique_multimer_complex_ids = unique(processed_multimer_output$complex_id)
# plot_complex_id = unique_multimer_complex_ids[16] # custom

non_AA_complex_ids = c("5oxz_A_5oxz_B", "6h7a_A_6h7a_B", "6tri_A_6tri_B", "7f9h_A_7f9h_B",
                       "7jqd_A_7jqd_B", "7n1j_A_7n1j_B", "7nto_A_7nto_B", "5ytq_A_5ytq_B",
                       "6otn_A_6otn_B", "6v7u_A_6v7u_B", "6dm9_A_6dm9_B")

AA_complex_ids = c("7phu_A_7phu_B", "7s0x_A_7s0x_B", "7sd5_A_7sd5_B", "7qez_A_7qez_B",
                   "8be1_A_8be1_B", "7mdp_A_7mdp_B")

# test_ids = c("5oxz_A_5oxz_B")

# plot_complex_id = "5oxz_A_5oxz_B"
  
save_heat_map = function(plot_complex_id, is_AA) {
  if (is_AA) {
    plot_complex_rows = processed_AA_multimer_output %>%
      filter(complex_id == plot_complex_id)
  } else {
    plot_complex_rows = processed_multimer_output %>%
      filter(complex_id == plot_complex_id)
  }
  
  # print(dim(plot_complex_rows))
  
  plot_seq1 = plot_complex_rows$seq1[1]
  plot_seq2 = plot_complex_rows$seq2[1]
  # print(plot_seq1)
  # print(plot_seq2)
  plot_seq = paste0(plot_seq1, ":", plot_seq2)
  
  # print(dim(plot_complex_rows))
  
  n_shift = nchar(plot_seq1)
  
  plot_data = master_simple_data_longer %>%
    filter(seq == plot_seq1 | seq == plot_seq2,
           delete_index >= 0) %>%
    mutate(delete_index = ifelse(seq == plot_seq1, delete_index, delete_index + n_shift),
           delta_score = ifelse(score_type == 'mon_delta_mean_pLDDT_score' | 
                                  score_type == 'multi_delta_mean_pLDDT_score',
                                delta_score/100.0,
                                delta_score),
           seq = paste0(plot_seq1, plot_seq2),
           seq_type = ifelse(score_type == 'mon_delta_mean_pLDDT_score' |
                                  score_type == 'mon_delta_pTM_score',
                             'Monomer',
                             'Multimer'),
           score_type = case_when(
             score_type == 'mon_delta_mean_pLDDT_score' ~ 'delta_mean_pLDDT_score',
             score_type == 'mon_delta_pTM_score' ~ 'delta_pTM_score',
             score_type == 'multi_delta_mean_pLDDT_score' ~ 'delta_mean_pLDDT_score',
             score_type == 'multi_delta_pTM_score' ~ 'delta_pTM_score',
             score_type == 'multi_delta_ipTM_score' ~ 'delta_ipTM_score',
           ),
           ) %>%
    arrange(delete_index)
  
  # print(dim(plot_data))
  
  # --- create desired heatmap --- #
  g_example_heatmap <-ggplot(plot_data, aes(x=delete_index, y=score_type, fill=delta_score)) +
    geom_tile(color= "white", size=0.1) +
    facet_wrap(~seq_type, nrow = 2, scales="free") +
    scale_fill_viridis(name="Delta Score", option ="A") +
    labs(title = "Delta scores by index of protein complex",
         subtitle = paste0("Protein complex id: ", plot_complex_id)) +
    xlab("Index") +
    ylab("Score Type") 
  
  # - save - #
  # plot_file_name = paste0("../analysis_plots/", plot_complex_id, "_delta_scores.png")
  
  if (is_AA) {
    plot_file_name = paste0("../analysis_plots/AA/", plot_complex_id, "_delta_scores.png")
  } else {
    plot_file_name = paste0("../analysis_plots/non_AA/", plot_complex_id, "_delta_scores.png")
  }
  
  ggsave(plot = g_example_heatmap, width = 9, height = 6, dpi = 300, filename = plot_file_name)
  
  0
}


# run and save
# for (complex_id in non_AA_complex_ids) {
#   save_heat_map(complex_id, FALSE)
# }

for (complex_id in AA_complex_ids) {
  save_heat_map(complex_id, TRUE)
}


```




